Multi-level modelling with Openflexo/FML - A contribution to the \mpc

### 2021-05-26 6-ModelPresentation.tex
Updates from Overleaf

\todo[inline]{phrase d'intro pour expliquer que les req étaient couverts au fur et à mesure (côté systématique)}

@@

\begin{lstlisting}
concept ActorType extends Type {
  ActorType[0..*] parentActorTypes;
  ...
  boolean isOrSpecializes(ActorType actorType) {    
    if (actorType == this)   
      return true;      
    for (p : parentActorTypes) {      
      if (p.isOrSpecializes(actorType)
        return true;        
      return false;    
    }
 ...
}

concept Actor extends MultiInstance {
 ...
 boolean hasActorType(ActorType actType) {      
  for (type : types) {
   if (type.isOrSpecializes(actType)) {
    return true;
   }
  }     
 }  
 ...
}
\end{lstlisting}

@@

\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
concept CodingTaskType extends SETaskType 
{
  ProgrammingLanguage[1..*] languages;
  ...
  public Coding newTask(String name) { 
    return new Coding(name,this); 
  }
}    
\end{lstlisting}

@@

\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
forEach (artifactType : types) {
    assert !(artifactType instanceof CodeArtifactType) | artifactType.implements(languages)
}
\end{lstlisting}

@@

\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
concept CodingTaskType extends SETaskType 
{
  ...
  public Coding newTask(String name) {
    Coding returned = new Coding(name,this); 
    if (languages.contains(ProgrammingLanguage.COBOL)) {
        returned.addToPerformingActors(getActor("Ann Smith"));
    }
    return returned;
  }
}    
\end{lstlisting}

\textbf{S8} requirement is simply captured with the definition of \textit{Testing} instance of \textit{SETaskType}, \textit{Tester} instance of \textit{ActorType}, and \textit{Test report} instance of \textit{ArtifactType}.
% TODO S9

@@

% TODO S13

@@

>>>>

The presentation of our solution to the \mpc follows a systematic methodology where modeling choices are introduced step by step, as the requirements are stated. Satisfaction of requirements are explained throughout this presentation.

%\todo[inline]{phrase d'intro pour expliquer que les req étaient couverts au fur et à mesure (côté systématique)}

@@

\begin{lstlisting}
concept ActorType extends Type {
  ActorType[0..*] parentActorTypes;
  ...
  boolean isOrSpecializes(ActorType actorType) {    
    if (actorType == this)   
      return true;      
    for (p : parentActorTypes) {      
      if (p.isOrSpecializes(actorType)
        return true;
    }
    return false;
  }
  ...
}

concept Actor extends MultiInstance {
  ...
  boolean hasActorType(ActorType actType) {      
    for (type : types) {
      if (type.isOrSpecializes(actType))
        return true;
    }
    return false;
  }     
  ...
}
\end{lstlisting}

@@

\begin{lstlisting}
concept CodingTaskType extends SETaskType {
  ProgrammingLanguage[1..*] languages;
  ...
  public Coding newTask(String name) { 
    return new Coding(name,this); 
  }
}    
\end{lstlisting}

@@

\begin{lstlisting}
forEach (artifactType : types) {
  assert !(artifactType instanceof CodeArtifactType) | artifactType.doesImplement(languages)
}
\end{lstlisting}

@@

\begin{lstlisting}
concept CodingTaskType extends SETaskType {
  ...
  public Coding newTask(String name) {
    Coding returned = new Coding(name,this); 
    if (languages.contains(ProgrammingLanguage.COBOL))
      returned.addToPerformingActors(getActor("Ann Smith"));
    return returned;
  }
}    
\end{lstlisting}

\textbf{S8} requirement is simply captured with the definition of \textit{Testing} instance of \textit{SETaskType}, \textit{Tester} instance of \textit{ActorType}, and \textit{Test report} instance of \textit{ArtifactType}.

\todo[inline]{requirement S9 non couvert pour le moment: compléter}

@@

\todo[inline]{requirement S13 pas complètement couvert : compléter}

@@

Our solution is fully implemented within Openflexo tool. Both usecases have been modelled in the interactive design environment and are executable by the FML execution engine.

\todo[inline]{mettre ici le lien vers la page web qui montre la démo}

Ici: un paragraphe qui décrit l'exécution, l'activation et la finalisation des taches, la business logic des gateways, etc...

We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming TA} to implement two interactive graphical tools built on top of conceptual levels detailed in previous section. A first tool offers a graphical edition of a Process type, while the second tool offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to some actors, and the execution of this process with a graphical visualization.

Figure \ref{fig:ToolingArchitecture} shows the architecture of these two tools. Left of figure presents process type graphical editor. ProcessTypeEditor is implemented as a VirtualModel declaring two model slots (represented with bold circles). The first model slot references an instance of Acme metamodel, while the second model slot references an instance of diagram, conforms to the ProcessType diagram specification (a diagram metamodel
